---
title: "DADS_compiled_figures"
output: html_document
date: "2024-06-11"
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/DADS")

library(phyloseq); packageVersion("phyloseq")
library(ggplot2); packageVersion("ggplot2")
library(readr); packageVersion("readr")
library(purrr); packageVersion("purrr")
library(furrr); packageVersion("furrr")
library(dplyr); packageVersion("dplyr")
library(stringr); packageVersion("stringr")
library(metacoder); packageVersion("metacoder")
library(data.table); packageVersion("data.table")
library(Biostrings); packageVersion("Biostrings")
library(vegan);packageVersion("vegan")
library(agricolae);packageVersion("agricolae")
library(magrittr);packageVersion("magrittr")
library(data.table);packageVersion("data.table")
library(DESeq2);packageVersion("DESeq2")
library(cowplot);packageVersion("cowplot")
library(doParallel);packageVersion("doParallel")
library(showtext)
font_add("Arial", regular = "path/to/arial.ttf")
showtext_auto()
```

### Load data from all three experiments
```{r subset data ITS, include=TRUE}
load("Data/ITS/filtered_data/dadsexp.RData")
dads_its.ps<-dads.ps

load("Data/ITS/filtered_data/dads_5asv.RData")
dads_5asv_its.ps<-dads_5asv.ps

load("Data/16S/filtered_data/dadsexp.RData")
dads_16S.ps<-dads.ps

load("Data/16S/filtered_data/dads_5asv.RData")
dads_5asv_16S.ps<-dads_5asv.ps

load("Data/RPS10/filtered_data/dadsexp.RData")
dads_rps10.ps<-dads.ps

load("Data/RPS10/filtered_data/dads_5asv.RData")
dads_5asv_rps10.ps<-dads_5asv.ps
```

### Let's calculate alpha diversity metrics for ITS dataset
```{r, combine alpha diversity boxplots its}
set.seed(1)
variables <- c("Treatment", "Soil", "Incubation")
measures <- c("Observed", "InvSimpson")

AD_its <- phyloseq::estimate_richness(dads_its.ps, measures = c("Observed", "InvSimpson"))

AD_its <- cbind(AD_its, sample_data(dads_its.ps)[rownames(AD_its), c("Treatment", "Soil", "Incubation")])

variables <- c("Treatment", "Soil", "Incubation")

plots_its <- list()

for (variable in variables) {
  p <- phyloseq::plot_richness(dads_its.ps, x = variable, measures = c("Observed", "InvSimpson")) + 
    geom_boxplot() +
    theme_classic() +  # Use theme_classic() as a starting point
    theme(
      panel.border = element_rect(color = "black", fill = NA, size = 1),  # Add border around plot panel
      axis.line = element_blank(),  # Remove axis lines (they'll be replaced by the panel border)
      plot.margin = unit(c(1, 1, 1, 1), "lines"),  # Add some margin around the plot
      plot.title = element_text(margin = margin(b = 10)),  # Add some space below the title
      axis.title.x = element_text(margin = margin(t = 10)),  # Add some space above the x-axis title
      axis.title.y = element_text(margin = margin(r = 10))  # Add some space to the right of the y-axis title
    )
  plots_its[[variable]] <- p
}

#Let's calculate Tukey results
# Function to perform Tukey HSD and extract letters
get_tukey_letters <- function(data, measure, variable) {
  formula <- as.formula(paste(measure, "~", variable))
  aov_result <- aov(formula, data = data)
  tukey_result <- HSD.test(aov_result, variable, group = TRUE)
  letters <- tukey_result$groups$groups
  return(letters)
}

# Perform Tukey HSD tests and store results
tukey_results_ITS <- list()
for (variable in variables) {
  for (measure in measures) {
    tryCatch({
      letters <- get_tukey_letters(AD_its, measure, variable)
      tukey_results_ITS[[paste(measure, variable, sep = "_")]] <- letters
    }, error = function(e) {
      message("Error in Tukey test for ", measure, " by ", variable, ": ", e$message)
    })
  }
}


print(tukey_results_ITS)
```

### Let's calculate alpha diversity metrics for 16S dataset
```{r, combine alpha diversity boxplots 16S}
set.seed(1)
variables <- c("Treatment", "Soil", "Incubation")
measures <- c("Observed", "InvSimpson")

AD_16S <- phyloseq::estimate_richness(dads_16S.ps, measures = c("Observed", "InvSimpson"))
rownames(AD_16S) <- gsub("\\X", "", rownames(AD_16S))


# Combine richness estimates with sample metadata
AD_16S <- cbind(AD_16S, sample_data(dads_16S.ps)[rownames(AD_16S), c("Treatment", "Soil", "Incubation")])

variables <- c("Treatment", "Soil", "Incubation")

# Create an empty list to store plots
plots_16S <- list()

# Loop through each variable and create the plot
for (variable in variables) {
  p <- phyloseq::plot_richness(dads_16S.ps, x = variable, measures = c("Observed", "InvSimpson")) + 
    geom_boxplot() +
    theme_classic() +  # Use theme_classic() as a starting point
    theme(
      panel.border = element_rect(color = "black", fill = NA, size = 1),  # Add border around plot panel
      axis.line = element_blank(),  # Remove axis lines (they'll be replaced by the panel border)
      plot.margin = unit(c(1, 1, 1, 1), "lines"),  # Add some margin around the plot
      plot.title = element_text(margin = margin(b = 10)),  # Add some space below the title
      axis.title.x = element_text(margin = margin(t = 10)),  # Add some space above the x-axis title
      axis.title.y = element_text(margin = margin(r = 10))  # Add some space to the right of the y-axis title
    )
  plots_16S[[variable]] <- p
}

#Let's calculate Tukey results
# Function to perform Tukey HSD and extract letters
get_tukey_letters <- function(data, measure, variable) {
  formula <- as.formula(paste(measure, "~", variable))
  aov_result <- aov(formula, data = data)
  tukey_result <- HSD.test(aov_result, variable, group = TRUE)
  letters <- tukey_result$groups$groups
  return(letters)
}

# Perform Tukey HSD tests and store results
tukey_results_16S <- list()
for (variable in variables) {
  for (measure in measures) {
    tryCatch({
      letters <- get_tukey_letters(AD_16S, measure, variable)
      tukey_results_16S[[paste(measure, variable, sep = "_")]] <- letters
    }, error = function(e) {
      message("Error in Tukey test for ", measure, " by ", variable, ": ", e$message)
    })
  }
}


print(tukey_results_16S)
```

### Let's calculate alpha diversity metrics for RPS10 dataset
```{r, combine alpha diversity boxplots rps10}
set.seed(1)

variables <- c("Treatment", "Soil", "Incubation")
measures <- c("Observed", "InvSimpson")

# Estimate alpha diversity
AD_RPS10 <- estimate_richness(dads_rps10.ps, measures = c("Observed", "InvSimpson"))

# Combine richness estimates with sample metadata
AD_RPS10 <- cbind(AD_RPS10, sample_data(dads_rps10.ps)[rownames(AD_RPS10), c("Treatment", "Soil", "Incubation")])

variables <- c("Treatment", "Soil", "Incubation")

# Create an empty list to store plots
plots_rps10 <- list()

# Loop through each variable and create the plot
for (variable in variables) {
  p <- phyloseq::plot_richness(dads_rps10.ps, x = variable, measures = c("Observed", "InvSimpson")) + 
    geom_boxplot() +
    theme_classic() +  # Use theme_classic() as a starting point
    theme(
      panel.border = element_rect(color = "black", fill = NA, size = 1),  # Add border around plot panel
      axis.line = element_blank(),  # Remove axis lines (they'll be replaced by the panel border)
      plot.margin = unit(c(1, 1, 1, 1), "lines"),  # Add some margin around the plot
      plot.title = element_text(margin = margin(b = 10)),  # Add some space below the title
      axis.title.x = element_text(margin = margin(t = 10)),  # Add some space above the x-axis title
      axis.title.y = element_text(margin = margin(r = 10))  # Add some space to the right of the y-axis title
    )
  plots_rps10[[variable]] <- p
}

#Let's calculate Tukey results
# Function to perform Tukey HSD and extract letters
get_tukey_letters <- function(data, measure, variable) {
  formula <- as.formula(paste(measure, "~", variable))
  aov_result <- aov(formula, data = data)
  tukey_result <- HSD.test(aov_result, variable, group = TRUE)
  letters <- tukey_result$groups$groups
  return(letters)
}

# Perform Tukey HSD tests and store results
tukey_results_RPS10 <- list()
for (variable in variables) {
  for (measure in measures) {
    tryCatch({
      letters <- get_tukey_letters(AD_RPS10, measure, variable)
      tukey_results_RPS10[[paste(measure, variable, sep = "_")]] <- letters
    }, error = function(e) {
      message("Error in Tukey test for ", measure, " by ", variable, ": ", e$message)
    })
  }
}


print(tukey_results_RPS10)
```

### Let's combine the treatment plots
```{r, combine all alpha trt box plots}
create_label <- function(text, x = 0.5, y = 0.95) {  # Centered the label horizontally
  ggdraw() + 
    draw_label(text, fontface = 'bold', x = x, y = y, hjust = 0.5, vjust = 1)  # Changed hjust to 0.5 for center alignment
}

label_A <- create_label("A.")
label_B <- create_label("B.")
label_C <- create_label("C.")

# Arrange plots
plots_arrange_its <- gridExtra::grid.arrange(
  plots_its[["Treatment"]]
)
plots_arrange_16S <- gridExtra::grid.arrange(
  plots_16S[["Treatment"]]
)
plots_arrange_rps10 <- gridExtra::grid.arrange(
  plots_rps10[["Treatment"]]
)

combined_plots_16S <- plot_grid(label_A, plots_arrange_16S, ncol = 2, rel_widths = c(0.1, 1))  # Reduced label width
combined_plots_its <- plot_grid(label_B, plots_arrange_its, ncol = 2, rel_widths = c(0.1, 1))  # Reduced label width
combined_plots_rps10 <- plot_grid(label_C, plots_arrange_rps10, ncol = 2, rel_widths = c(0.1, 1))  # Reduced label width

facetted_figure_trt <- plot_grid(
  combined_plots_16S,
  combined_plots_its,
  combined_plots_rps10,
  nrow = 3,
  align = 'v',
  axis = 'l'
)


facetted_figure_trt_with_margins <- plot_grid(
  NULL, facetted_figure_trt, NULL,
  ncol = 3,
  rel_widths = c(0, 1, 0.08)  # 5% margin on each side
)

facetted_figure_trt_with_margins
# Save the figure
ggsave("Figures/facetted_figure_alphadiv_treatment.svg", facetted_figure_trt_with_margins, width=8, height=12)
ggsave("Figures/facetted_figure_alphadiv_treatment.pdf", facetted_figure_trt_with_margins, width=8, height=12)
```

### Let's combine the soil type plots
```{r, combine all alpha soil box plots}
create_label <- function(text, x = 0.5, y = 0.95) {  # Centered the label horizontally
  ggdraw() + 
    draw_label(text, fontface = 'bold', x = x, y = y, hjust = 0.5, vjust = 1)  # Changed hjust to 0.5 for center alignment
}

label_A <- create_label("A.")
label_B <- create_label("B.")
label_C <- create_label("C.")

# Arrange plots
plots_arrange_its <- gridExtra::grid.arrange(
  plots_its[["Soil"]]
)
plots_arrange_16S <- gridExtra::grid.arrange(
  plots_16S[["Soil"]]
)
plots_arrange_rps10 <- gridExtra::grid.arrange(
  plots_rps10[["Soil"]]
)

combined_plots_16S <- plot_grid(label_A, plots_arrange_16S, ncol = 2, rel_widths = c(0.1, 1))  # Reduced label width
combined_plots_its <- plot_grid(label_B, plots_arrange_its, ncol = 2, rel_widths = c(0.1, 1))  # Reduced label width
combined_plots_rps10 <- plot_grid(label_C, plots_arrange_rps10, ncol = 2, rel_widths = c(0.1, 1))  # Reduced label width

facetted_figure_soil <- plot_grid(
  combined_plots_16S,
  combined_plots_its,
  combined_plots_rps10,
  nrow = 3,
  align = 'v',
  axis = 'l'
)


facetted_figure_soil_with_margins <- plot_grid(
  NULL, facetted_figure_soil, NULL,
  ncol = 3,
  rel_widths = c(0, 1, 0.08)  # 5% margin on each side
)

facetted_figure_soil_with_margins 
# Save the figure
ggsave("Figures/facetted_figure_alphadiv_soil.svg", facetted_figure_soil_with_margins, width=8, height=12)
ggsave("Figures/facetted_figure_alphadiv_soil.pdf", facetted_figure_soil_with_margins, width=8, height=12)
```

### Let's combine the incubation plots
```{r, combine all alpha incubation box plots}
create_label <- function(text, x = 0.5, y = 0.95) {  # Centered the label horizontally
  ggdraw() + 
    draw_label(text, fontface = 'bold', x = x, y = y, hjust = 0.5, vjust = 1)  # Changed hjust to 0.5 for center alignment
}

label_A <- create_label("A.")
label_B <- create_label("B.")
label_C <- create_label("C.")

# Arrange plots
plots_arrange_its <- gridExtra::grid.arrange(
  plots_its[["Incubation"]]
)
plots_arrange_16S <- gridExtra::grid.arrange(
  plots_16S[["Incubation"]]
)
plots_arrange_rps10 <- gridExtra::grid.arrange(
  plots_rps10[["Incubation"]]
)

combined_plots_16S <- plot_grid(label_A, plots_arrange_16S, ncol = 2, rel_widths = c(0.1, 1))  # Reduced label width
combined_plots_its <- plot_grid(label_B, plots_arrange_its, ncol = 2, rel_widths = c(0.1, 1))  # Reduced label width
combined_plots_rps10 <- plot_grid(label_C, plots_arrange_rps10, ncol = 2, rel_widths = c(0.1, 1))  # Reduced label width

facetted_figure_inc <- plot_grid(
  combined_plots_16S,
  combined_plots_its,
  combined_plots_rps10,
  nrow = 3,
  align = 'v',
  axis = 'l'
)


facetted_figure_inc_with_margins <- plot_grid(
  NULL, facetted_figure_inc, NULL,
  ncol = 3,
  rel_widths = c(0, 1, 0.08)  # 5% margin on each side
)

facetted_figure_inc_with_margins 
# Save the figure
ggsave("Figures/facetted_figure_alphadiv_incubation.svg", facetted_figure_inc_with_margins, width=8, height=12)
ggsave("Figures/facetted_figure_alphadiv_incubation.pdf", facetted_figure_inc_with_margins, width=8, height=12)
```

### Let's proceed with ordination for the ITS dataset-PCoA
```{r, ordination plots its PCOA}
set.seed(1)
strat_rel_dads_its.ps<-transform_sample_counts(dads_5asv_its.ps, function(x) x / sum(x))

ps_PCoA_its = ordinate(strat_rel_dads_its.ps, 
                   method="PCoA", 
                   distance="bray")

relative_eig <- ps_PCoA_its$values$Relative_eig

ordination_df <- data.frame(ps_PCoA_its$vectors, sample_data(strat_rel_dads_its.ps))

plot_its_trt <- ggplot(ordination_df, aes(x = Axis.1, y = Axis.2, fill = Treatment)) +
  geom_point(size = 4, shape = 21, color = "black", stroke = 0.5) +
  labs(x = paste0("Axis 1 [", round(relative_eig[1] * 100, 1), "%]"), 
       y = paste0("Axis 2 [", round(relative_eig[2] * 100, 1), "%]"), 
       fill = "Treatment") +
  scale_x_continuous(breaks = seq(-0.5, 0.5, by = 0.5), limits = c(-0.5, 0.5)) +
  scale_y_continuous(breaks = seq(-0.5, 0.5, by = 0.5), limits = c(-0.5, 0.5)) +
  guides(fill = guide_legend(nrow = 2)) +
  theme_bw() +
  theme(axis.text = element_text(size = 14, family = "Arial"), 
        axis.title = element_text(size = 14, family = "Arial"), 
        title = element_text(size = 14, family = "Arial"), 
        legend.text = element_text(size = 14, family = "Arial"),
        legend.position = "bottom", 
        legend.margin = margin(0, 0, 0, 0),  
        legend.box.margin = margin(0, 0, 0, 0),  
        legend.key.size = unit(0.4, "lines"), 
        legend.spacing.x = unit(0.2, "cm"),   
        legend.spacing.y = unit(0.2, "cm"))


print(plot_its_trt)

plot_its_inc <- ggplot(ordination_df, aes(x = Axis.1, y = Axis.2, fill = Incubation)) +
  geom_point(size = 4, shape = 21, color = "black", stroke = 0.5) +
  labs(x = paste0("Axis 1 [", round(relative_eig[1] * 100, 1), "%]"), 
       y = paste0("Axis 2 [", round(relative_eig[2] * 100, 1), "%]"), 
       fill = "Incubation") +
  scale_x_continuous(breaks = seq(-0.5, 0.5, by = 0.5), limits = c(-0.5, 0.5)) +
  scale_y_continuous(breaks = seq(-0.5, 0.5, by = 0.5), limits = c(-0.5, 0.5)) +
  guides(fill = guide_legend(nrow = 2)) +
  theme_bw() +
  theme(axis.text = element_text(size = 14, family = "Arial"), 
        axis.title = element_text(size = 14, family = "Arial"), 
        title = element_text(size = 14, family = "Arial"), 
        legend.text = element_text(size = 14, family = "Arial"),
        legend.position = "bottom", 
        legend.margin = margin(0, 0, 0, 0),  
        legend.box.margin = margin(0, 0, 0, 0),  
        legend.key.size = unit(0.4, "lines"), 
        legend.spacing.x = unit(0.2, "cm"),   
        legend.spacing.y = unit(0.2, "cm"))

print(plot_its_inc)


plot_its_soil <- ggplot(ordination_df, aes(x = Axis.1, y = Axis.2, fill = Soil)) +
  geom_point(size = 4, shape = 21, color = "black", stroke = 0.5) +
  labs(x = paste0("Axis 1 [", round(relative_eig[1] * 100, 1), "%]"), 
       y = paste0("Axis 2 [", round(relative_eig[2] * 100, 1), "%]"), 
       fill = "Soil") +
  scale_x_continuous(breaks = seq(-0.5, 0.5, by = 0.5), limits = c(-0.5, 0.5)) +
  scale_y_continuous(breaks = seq(-0.5, 0.5, by = 0.5), limits = c(-0.5, 0.5)) +
  guides(fill = guide_legend(nrow = 2)) +
  theme_bw() +
  theme(axis.text = element_text(size = 14, family = "Arial"), 
        axis.title = element_text(size = 14, family = "Arial"), 
        title = element_text(size = 14, family = "Arial"), 
        legend.text = element_text(size = 14, family = "Arial"),
        legend.position = "bottom", 
        legend.margin = margin(0, 0, 0, 0),  
        legend.box.margin = margin(0, 0, 0, 0),  
        legend.key.size = unit(0.4, "lines"), 
        legend.spacing.x = unit(0.2, "cm"),   
        legend.spacing.y = unit(0.2, "cm"))

print(plot_its_soil)
```


### Let's proceed with ordination for the 16S dataset-PCoA
```{r, ordination plots 16s PCOA}
set.seed(1)
strat_rel_dads_16S.ps<-transform_sample_counts(dads_5asv_16S.ps, function(x) x / sum(x))

ps_PCoA_16S = ordinate(strat_rel_dads_16S.ps, 
                   method="PCoA", 
                   distance="bray")

relative_eig <- ps_PCoA_16S$values$Relative_eig

ordination_df <- data.frame(ps_PCoA_16S$vectors, sample_data(strat_rel_dads_16S.ps))

plot_16S_trt <- ggplot(ordination_df, aes(x = Axis.1, y = Axis.2, fill = Treatment)) +
  geom_point(size = 4, shape = 21, color = "black", stroke = 0.5) +
  labs(x = paste0("Axis 1 [", round(relative_eig[1] * 100, 1), "%]"), 
       y = paste0("Axis 2 [", round(relative_eig[2] * 100, 1), "%]"), 
       fill = "Treatment") +
  scale_x_continuous(breaks = seq(-0.5, 0.5, by = 0.5), limits = c(-0.5, 0.5)) +
  scale_y_continuous(breaks = seq(-0.5, 0.5, by = 0.5), limits = c(-0.5, 0.5)) +
  guides(fill = guide_legend(nrow = 2)) +
  theme_bw() +
  theme(axis.text = element_text(size = 14, family = "Arial"), 
        axis.title = element_text(size = 14, family = "Arial"), 
        title = element_text(size = 14, family = "Arial"), 
        legend.text = element_text(size = 14, family = "Arial"),
        legend.position = "bottom", 
        legend.margin = margin(0, 0, 0, 0),  
        legend.box.margin = margin(0, 0, 0, 0),  
        legend.key.size = unit(0.4, "lines"), 
        legend.spacing.x = unit(0.2, "cm"),   
        legend.spacing.y = unit(0.2, "cm"))


print(plot_16S_trt)

plot_16S_inc <- ggplot(ordination_df, aes(x = Axis.1, y = Axis.2, fill = Incubation)) +
  geom_point(size = 4, shape = 21, color = "black", stroke = 0.5) +
  labs(x = paste0("Axis 1 [", round(relative_eig[1] * 100, 1), "%]"), 
       y = paste0("Axis 2 [", round(relative_eig[2] * 100, 1), "%]"), 
       fill = "Incubation") +
  scale_x_continuous(breaks = seq(-0.5, 0.5, by = 0.5), limits = c(-0.5, 0.5)) +
  scale_y_continuous(breaks = seq(-0.5, 0.5, by = 0.5), limits = c(-0.5, 0.5)) +
  guides(fill = guide_legend(nrow = 2)) +
  theme_bw() +
  theme(axis.text = element_text(size = 14, family = "Arial"), 
        axis.title = element_text(size = 14, family = "Arial"), 
        title = element_text(size = 14, family = "Arial"), 
        legend.text = element_text(size = 14, family = "Arial"),
        legend.position = "bottom", 
        legend.margin = margin(0, 0, 0, 0),  
        legend.box.margin = margin(0, 0, 0, 0),  
        legend.key.size = unit(0.4, "lines"), 
        legend.spacing.x = unit(0.2, "cm"),   
        legend.spacing.y = unit(0.2, "cm"))

print(plot_16S_inc)


plot_16S_soil <- ggplot(ordination_df, aes(x = Axis.1, y = Axis.2, fill = Soil)) +
  geom_point(size = 4, shape = 21, color = "black", stroke = 0.5) +
  labs(x = paste0("Axis 1 [", round(relative_eig[1] * 100, 1), "%]"), 
       y = paste0("Axis 2 [", round(relative_eig[2] * 100, 1), "%]"), 
       fill = "Soil") +
  scale_x_continuous(breaks = seq(-0.5, 0.5, by = 0.5), limits = c(-0.5, 0.5)) +
  scale_y_continuous(breaks = seq(-0.5, 0.5, by = 0.5), limits = c(-0.5, 0.5)) +
  guides(fill = guide_legend(nrow = 2)) +
  theme_bw() +
  theme(axis.text = element_text(size = 14, family = "Arial"), 
        axis.title = element_text(size = 14, family = "Arial"), 
        title = element_text(size = 14, family = "Arial"), 
        legend.text = element_text(size = 14, family = "Arial"),
        legend.position = "bottom", 
        legend.margin = margin(0, 0, 0, 0),  
        legend.box.margin = margin(0, 0, 0, 0),  
        legend.key.size = unit(0.4, "lines"), 
        legend.spacing.x = unit(0.2, "cm"),   
        legend.spacing.y = unit(0.2, "cm"))

print(plot_16S_soil)
```

### Let's proceed with ordination for the RPS10 dataset-pCoA
```{r, ordination plots rps10 PCOA}
set.seed(1)
strat_rel_dads_rps10.ps<-transform_sample_counts(dads_5asv_rps10.ps, function(x) x / sum(x))

ps_PCoA_rps10 = ordinate(strat_rel_dads_rps10.ps, 
                   method="PCoA", 
                   distance="bray")

ordination_df <- data.frame(ps_PCoA_rps10$vectors, sample_data(strat_rel_dads_rps10.ps))

# Assuming Relative_eig values are stored in ps_PCoA_rps10$values$Relative_eig
relative_eig <- ps_PCoA_rps10$values$Relative_eig

# Modify the plot code to include Relative_eig values in the axis labels
plot_rps10_trt <- ggplot(ordination_df, aes(x = Axis.1, y = Axis.2, fill = Treatment)) +
  geom_point(size = 4, shape = 21, color = "black", stroke = 0.5) +
  labs(x = paste0("Axis 1 [", round(relative_eig[1] * 100, 1), "%]"), 
       y = paste0("Axis 2 [", round(relative_eig[2] * 100, 1), "%]"), 
       fill = "Treatment") +
  scale_x_continuous(breaks = seq(-0.5, 0.5, by = 0.5), limits = c(-0.5, 0.5)) +
  scale_y_continuous(breaks = seq(-0.5, 0.5, by = 0.5), limits = c(-0.5, 0.5)) +
  guides(fill = guide_legend(nrow = 2)) +
  theme_bw() +
  theme(axis.text = element_text(size = 14, family = "Arial"), 
        axis.title = element_text(size = 14, family = "Arial"), 
        title = element_text(size = 14, family = "Arial"), 
        legend.text = element_text(size = 14, family = "Arial"),
        legend.position = "bottom", 
        legend.margin = margin(0, 0, 0, 0),  
        legend.box.margin = margin(0, 0, 0, 0),  
        legend.key.size = unit(0.4, "lines"), 
        legend.spacing.x = unit(0.2, "cm"),   
        legend.spacing.y = unit(0.2, "cm"))

print(plot_rps10_trt)

plot_rps10_inc <- ggplot(ordination_df, aes(x = Axis.1, y = Axis.2, fill = Incubation)) +
  geom_point(size = 4, shape = 21, color = "black", stroke = 0.5) +
  labs(x = paste0("Axis 1 [", round(relative_eig[1] * 100, 1), "%]"), 
       y = paste0("Axis 2 [", round(relative_eig[2] * 100, 1), "%]"), 
       fill = "Incubation") +
  scale_x_continuous(breaks = seq(-0.5, 0.5, by = 0.5), limits = c(-0.5, 0.5)) +
  scale_y_continuous(breaks = seq(-0.5, 0.5, by = 0.5), limits = c(-0.5, 0.5)) +
  guides(fill = guide_legend(nrow = 2)) +
  theme_bw() +
  theme(axis.text = element_text(size = 14, family = "Arial"), 
        axis.title = element_text(size = 14, family = "Arial"), 
        title = element_text(size = 14, family = "Arial"), 
        legend.text = element_text(size = 14, family = "Arial"),
        legend.position = "bottom", 
        legend.margin = margin(0, 0, 0, 0),  
        legend.box.margin = margin(0, 0, 0, 0),  
        legend.key.size = unit(0.4, "lines"), 
        legend.spacing.x = unit(0.2, "cm"),   
        legend.spacing.y = unit(0.2, "cm"))

print(plot_rps10_inc)


plot_rps10_soil <- ggplot(ordination_df, aes(x = Axis.1, y = Axis.2, fill = Soil)) +
  geom_point(size = 4, shape = 21, color = "black", stroke = 0.5) +
  labs(x = paste0("Axis 1 [", round(relative_eig[1] * 100, 1), "%]"), 
       y = paste0("Axis 2 [", round(relative_eig[2] * 100, 1), "%]"), 
       fill = "Soil") +
  scale_x_continuous(breaks = seq(-0.5, 0.5, by = 0.5), limits = c(-0.5, 0.5)) +
  scale_y_continuous(breaks = seq(-0.5, 0.5, by = 0.5), limits = c(-0.5, 0.5)) +
  guides(fill = guide_legend(nrow = 2)) +
  theme_bw() +
  theme(axis.text = element_text(size = 14, family = "Arial"), 
        axis.title = element_text(size = 14, family = "Arial"), 
        title = element_text(size = 14, family = "Arial"), 
        legend.text = element_text(size = 14, family = "Arial"),
        legend.position = "bottom", 
        legend.margin = margin(0, 0, 0, 0),  
        legend.box.margin = margin(0, 0, 0, 0),  
        legend.key.size = unit(0.4, "lines"), 
        legend.spacing.x = unit(0.2, "cm"),   
        legend.spacing.y = unit(0.2, "cm"))

print(plot_rps10_soil)
```

### Let's combine plots into facetted plot
```{r, box plots ordination pcoa}
library(ggplot2)
library(cowplot)

create_column <- function(ord_plot) {
  plot_grid(
    ord_plot,
    nrow = 1,
    align = 'v',
    axis = 'l',
    rel_heights = c(1)
  )
}
# Create columns for each dataset and factor with adjusted spacing
col_16S_treatment <- create_column(plot_16S_trt)
col_16S_soil <- create_column(plot_16S_soil)
col_16S_incubation <- create_column(plot_16S_inc)
col_its_treatment <- create_column(plot_its_trt)
col_its_soil <- create_column(plot_its_soil)
col_its_incubation <- create_column(plot_its_inc)
col_rps10_treatment <- create_column(plot_rps10_trt)
col_rps10_soil <- create_column(plot_rps10_soil)
col_rps10_incubation <- create_column(plot_rps10_inc)

# Create row plots with labels and adjusted spacing
row_16S <- plot_grid(col_16S_treatment, col_16S_soil, col_16S_incubation, 
                     ncol = 3, labels = "A.", label_size = 20, label_x =  -0.01, label_y = 1.02, hjust = -0.5, vjust = 2, rel_widths = c(1, 1, 1))
row_its <- plot_grid(col_its_treatment, col_its_soil, col_its_incubation, 
                     ncol = 3, labels = "B.", label_size = 20, label_x =  -0.01, label_y = 1.02, hjust = -0.5, vjust = 2, rel_widths = c(1, 1, 1))
row_rps10 <- plot_grid(col_rps10_treatment, col_rps10_soil, col_rps10_incubation, 
                       ncol = 3, labels = "C.", label_size = 20, label_x =  -0.01, label_y = 1.02, hjust = -0.5, vjust = 2, rel_widths = c(1, 1, 1))


# Combine all elements with adjusted spacing
facetted_figure <- plot_grid(
  row_16S,
  row_its,
  row_rps10,
  nrow = 3,
  align = 'v',
  axis = 'l',
  rel_heights = c(1, 1, 1)  # Adjust relative heights
)

library(extrafont)
# Save the figure with increased dimensions
ggsave("Figures/facetted_figure_with_ordination_PCoA.svg", facetted_figure, width = 14, height = 14, units = "in")
ggsave("Figures/facetted_figure_with_ordination_PCoA.pdf", facetted_figure, width = 14, height = 14, units = "in")
```

### Let's proceed with ordination for the ITS dataset-NMDS
```{r, ordination plots its NMDS}
set.seed(1)
strat_rel_dads_its.ps<-transform_sample_counts(dads_5asv_its.ps, function(x) x / sum(x))

ps_NMDS_its = ordinate(strat_rel_dads_its.ps, 
                   method="NMDS", 
                   distance="bray")

relative_eig <- ps_NMDS_its$values$Relative_eig

ordination_df <- data.frame(ps_NMDS_its$points, sample_data(strat_rel_dads_its.ps))

plot_its_trt <- ggplot(ordination_df, aes(x = MDS1, y = MDS2, fill = Treatment)) +
  geom_point(size = 4, shape = 21, color = "black", stroke = 0.5, position = position_jitter(width = 0.05, height = 0.05)) +
  labs(x = "NMDS1",
       y = "NMDS2", 
       fill = "Treatment") +
  #scale_x_continuous(breaks = seq(-1, 1, by = 0.5), limits = c(-1, 1)) +
  #scale_y_continuous(breaks = seq(-1, 1, by = 0.5), limits = c(-1, 1)) +
  guides(fill = guide_legend(nrow = 2)) +
  theme_bw() +
  theme(axis.text = element_text(size = 14, family = "Arial"), 
        axis.title = element_text(size = 14, family = "Arial"), 
        title = element_text(size = 14, family = "Arial"), 
        legend.text = element_text(size = 14, family = "Arial"),
        legend.position = "bottom", 
        legend.margin = margin(0, 0, 0, 0),  
        legend.box.margin = margin(0, 0, 0, 0),  
        legend.key.size = unit(0.4, "lines"), 
        legend.spacing.x = unit(0.2, "cm"),   
        legend.spacing.y = unit(0.2, "cm"))


print(plot_its_trt)

plot_its_inc <- ggplot(ordination_df, aes(x = MDS1, y = MDS2, fill = Incubation)) +
  geom_point(size = 4, shape = 21, color = "black", stroke = 0.5, position = position_jitter(width = 0.05, height = 0.05)) +
  labs(x = "NMDS1",
       y = "NMDS2", 
       fill = "Incubation") +
  #scale_x_continuous(breaks = seq(-0.5, 0.5, by = 0.5), limits = c(-0.5, 0.5)) +
  #scale_y_continuous(breaks = seq(-0.5, 0.5, by = 0.5), limits = c(-0.5, 0.5)) +
  guides(fill = guide_legend(nrow = 2)) +
  theme_bw() +
  theme(axis.text = element_text(size = 14, family = "Arial"), 
        axis.title = element_text(size = 14, family = "Arial"), 
        title = element_text(size = 14, family = "Arial"), 
        legend.text = element_text(size = 14, family = "Arial"),
        legend.position = "bottom", 
        legend.margin = margin(0, 0, 0, 0),  
        legend.box.margin = margin(0, 0, 0, 0),  
        legend.key.size = unit(0.4, "lines"), 
        legend.spacing.x = unit(0.2, "cm"),   
        legend.spacing.y = unit(0.2, "cm"))

print(plot_its_inc)


plot_its_soil <- ggplot(ordination_df, aes(x = MDS1, y = MDS2, fill = Soil)) +
  geom_point(size = 4, shape = 21, color = "black", stroke = 0.5, position = position_jitter(width = 0.05, height = 0.05)) +
  labs(x = "NMDS1",
       y = "NMDS2", 
       fill = "Treatment") +
  #scale_x_continuous(breaks = seq(-0.5, 0.5, by = 0.5), limits = c(-0.5, 0.5)) +
  #scale_y_continuous(breaks = seq(-0.5, 0.5, by = 0.5), limits = c(-0.5, 0.5)) +
  guides(fill = guide_legend(nrow = 2)) +
  theme_bw() +
  theme(axis.text = element_text(size = 14, family = "Arial"), 
        axis.title = element_text(size = 14, family = "Arial"), 
        title = element_text(size = 14, family = "Arial"), 
        legend.text = element_text(size = 14, family = "Arial"),
        legend.position = "bottom", 
        legend.margin = margin(0, 0, 0, 0),  
        legend.box.margin = margin(0, 0, 0, 0),  
        legend.key.size = unit(0.4, "lines"), 
        legend.spacing.x = unit(0.2, "cm"),   
        legend.spacing.y = unit(0.2, "cm"))

print(plot_its_soil)
```


### Let's proceed with ordination for the 16S dataset-NMDS
```{r, ordination plots 16s NMDS}
set.seed(1)
strat_rel_dads_16S.ps<-transform_sample_counts(dads_5asv_16S.ps, function(x) x / sum(x))

ps_NMDS_16S = ordinate(strat_rel_dads_16S.ps, 
                   method="NMDS", 
                   distance="bray")

relative_eig <- ps_NMDS_16S$values$Relative_eig

ordination_df <- data.frame(ps_NMDS_16S$points, sample_data(strat_rel_dads_16S.ps))

plot_16S_trt <- ggplot(ordination_df, aes(x = MDS1, y = MDS2, fill = Treatment)) +
  geom_point(size = 4, shape = 21, color = "black", stroke = 0.5, position = position_jitter(width = 0.05, height = 0.05)) +
  labs(x = "NMDS1",
       y = "NMDS2", 
       fill = "Treatment") +
  #scale_x_continuous(breaks = seq(-1, 1, by = 0.5), lim16S = c(-1, 1)) +
  #scale_y_continuous(breaks = seq(-1, 1, by = 0.5), lim16S = c(-1, 1)) +
  guides(fill = guide_legend(nrow = 2)) +
  theme_bw() +
  theme(axis.text = element_text(size = 14, family = "Arial"), 
        axis.title = element_text(size = 14, family = "Arial"), 
        title = element_text(size = 14, family = "Arial"), 
        legend.text = element_text(size = 14, family = "Arial"),
        legend.position = "bottom", 
        legend.margin = margin(0, 0, 0, 0),  
        legend.box.margin = margin(0, 0, 0, 0),  
        legend.key.size = unit(0.4, "lines"), 
        legend.spacing.x = unit(0.2, "cm"),   
        legend.spacing.y = unit(0.2, "cm"))


print(plot_16S_trt)

plot_16S_inc <- ggplot(ordination_df, aes(x = MDS1, y = MDS2, fill = Incubation)) +
  geom_point(size = 4, shape = 21, color = "black", stroke = 0.5, position = position_jitter(width = 0.05, height = 0.05)) +
  labs(x = "NMDS1",
       y = "NMDS2", 
       fill = "Incubation") +
  #scale_x_continuous(breaks = seq(-0.5, 0.5, by = 0.5), lim16S = c(-0.5, 0.5)) +
  #scale_y_continuous(breaks = seq(-0.5, 0.5, by = 0.5), lim16S = c(-0.5, 0.5)) +
  guides(fill = guide_legend(nrow = 2)) +
  theme_bw() +
  theme(axis.text = element_text(size = 14, family = "Arial"), 
        axis.title = element_text(size = 14, family = "Arial"), 
        title = element_text(size = 14, family = "Arial"), 
        legend.text = element_text(size = 14, family = "Arial"),
        legend.position = "bottom", 
        legend.margin = margin(0, 0, 0, 0),  
        legend.box.margin = margin(0, 0, 0, 0),  
        legend.key.size = unit(0.4, "lines"), 
        legend.spacing.x = unit(0.2, "cm"),   
        legend.spacing.y = unit(0.2, "cm"))

print(plot_16S_inc)


plot_16S_soil <- ggplot(ordination_df, aes(x = MDS1, y = MDS2, fill = Soil)) +
  geom_point(size = 4, shape = 21, color = "black", stroke = 0.5, position = position_jitter(width = 0.05, height = 0.05)) +
  labs(x = "NMDS1",
       y = "NMDS2", 
       fill = "Treatment") +
  #scale_x_continuous(breaks = seq(-0.5, 0.5, by = 0.5), lim16S = c(-0.5, 0.5)) +
  #scale_y_continuous(breaks = seq(-0.5, 0.5, by = 0.5), lim16S = c(-0.5, 0.5)) +
  guides(fill = guide_legend(nrow = 2)) +
  theme_bw() +
  theme(axis.text = element_text(size = 14, family = "Arial"), 
        axis.title = element_text(size = 14, family = "Arial"), 
        title = element_text(size = 14, family = "Arial"), 
        legend.text = element_text(size = 14, family = "Arial"),
        legend.position = "bottom", 
        legend.margin = margin(0, 0, 0, 0),  
        legend.box.margin = margin(0, 0, 0, 0),  
        legend.key.size = unit(0.4, "lines"), 
        legend.spacing.x = unit(0.2, "cm"),   
        legend.spacing.y = unit(0.2, "cm"))

print(plot_16S_soil)
```

### Let's proceed with ordination for the RPS10 dataset-NMDS
```{r, ordination plots rps10 NMDS}
set.seed(1)
strat_rel_dads_rps10.ps<-transform_sample_counts(dads_5asv_rps10.ps, function(x) x / sum(x))

ps_NMDS_rps10 = ordinate(strat_rel_dads_rps10.ps, 
                   method="NMDS", 
                   distance="bray")

relative_eig <- ps_NMDS_rps10$values$Relative_eig

ordination_df <- data.frame(ps_NMDS_rps10$points, sample_data(strat_rel_dads_rps10.ps))

plot_rps10_trt <- ggplot(ordination_df, aes(x = MDS1, y = MDS2, fill = Treatment)) +
  geom_point(size = 4, shape = 21, color = "black", stroke = 0.5, position = position_jitter(width = 0.05, height = 0.05)) +
  labs(x = "NMDS1",
       y = "NMDS2", 
       fill = "Treatment") +
  #scale_x_continuous(breaks = seq(-1, 1, by = 0.5), limrps10 = c(-1, 1)) +
  #scale_y_continuous(breaks = seq(-1, 1, by = 0.5), limrps10 = c(-1, 1)) +
  guides(fill = guide_legend(nrow = 2)) +
  theme_bw() +
  theme(axis.text = element_text(size = 14, family = "Arial"), 
        axis.title = element_text(size = 14, family = "Arial"), 
        title = element_text(size = 14, family = "Arial"), 
        legend.text = element_text(size = 14, family = "Arial"),
        legend.position = "bottom", 
        legend.margin = margin(0, 0, 0, 0),  
        legend.box.margin = margin(0, 0, 0, 0),  
        legend.key.size = unit(0.4, "lines"), 
        legend.spacing.x = unit(0.2, "cm"),   
        legend.spacing.y = unit(0.2, "cm"))


print(plot_rps10_trt)

plot_rps10_inc <- ggplot(ordination_df, aes(x = MDS1, y = MDS2, fill = Incubation)) +
  geom_point(size = 4, shape = 21, color = "black", stroke = 0.5, position = position_jitter(width = 0.05, height = 0.05)) +
  labs(x = "NMDS1",
       y = "NMDS2", 
       fill = "Incubation") +
  #scale_x_continuous(breaks = seq(-0.5, 0.5, by = 0.5), limrps10 = c(-0.5, 0.5)) +
  #scale_y_continuous(breaks = seq(-0.5, 0.5, by = 0.5), limrps10 = c(-0.5, 0.5)) +
  guides(fill = guide_legend(nrow = 2)) +
  theme_bw() +
  theme(axis.text = element_text(size = 14, family = "Arial"), 
        axis.title = element_text(size = 14, family = "Arial"), 
        title = element_text(size = 14, family = "Arial"), 
        legend.text = element_text(size = 14, family = "Arial"),
        legend.position = "bottom", 
        legend.margin = margin(0, 0, 0, 0),  
        legend.box.margin = margin(0, 0, 0, 0),  
        legend.key.size = unit(0.4, "lines"), 
        legend.spacing.x = unit(0.2, "cm"),   
        legend.spacing.y = unit(0.2, "cm"))

print(plot_rps10_inc)


plot_rps10_soil <- ggplot(ordination_df, aes(x = MDS1, y = MDS2, fill = Soil)) +
  geom_point(size = 4, shape = 21, color = "black", stroke = 0.5, position = position_jitter(width = 0.05, height = 0.05)) +
  labs(x = "NMDS1",
       y = "NMDS2", 
       fill = "Treatment") +
  #scale_x_continuous(breaks = seq(-0.5, 0.5, by = 0.5), limrps10 = c(-0.5, 0.5)) +
  #scale_y_continuous(breaks = seq(-0.5, 0.5, by = 0.5), limrps10 = c(-0.5, 0.5)) +
  guides(fill = guide_legend(nrow = 2)) +
  theme_bw() +
  theme(axis.text = element_text(size = 14, family = "Arial"), 
        axis.title = element_text(size = 14, family = "Arial"), 
        title = element_text(size = 14, family = "Arial"), 
        legend.text = element_text(size = 14, family = "Arial"),
        legend.position = "bottom", 
        legend.margin = margin(0, 0, 0, 0),  
        legend.box.margin = margin(0, 0, 0, 0),  
        legend.key.size = unit(0.4, "lines"), 
        legend.spacing.x = unit(0.2, "cm"),   
        legend.spacing.y = unit(0.2, "cm"))

print(plot_rps10_soil)
```

### Let's combine plots into facetted plot
```{r, box plots ordination NMDS}
library(ggplot2)
library(cowplot)

create_column <- function(ord_plot) {
  plot_grid(
    ord_plot,
    nrow = 1,
    align = 'v',
    axis = 'l',
    rel_heights = c(1)
  )
}
# Create columns for each dataset and factor with adjusted spacing
col_16S_treatment <- create_column(plot_16S_trt)
col_16S_soil <- create_column(plot_16S_soil)
col_16S_incubation <- create_column(plot_16S_inc)
col_its_treatment <- create_column(plot_its_trt)
col_its_soil <- create_column(plot_its_soil)
col_its_incubation <- create_column(plot_its_inc)
col_rps10_treatment <- create_column(plot_rps10_trt)
col_rps10_soil <- create_column(plot_rps10_soil)
col_rps10_incubation <- create_column(plot_rps10_inc)

# Create row plots with labels and adjusted spacing
row_16S <- plot_grid(col_16S_treatment, col_16S_soil, col_16S_incubation, 
                     ncol = 3, labels = "A.", label_size = 20, label_x =  -0.01, label_y = 1.02, hjust = -0.5, vjust = 2, rel_widths = c(1, 1, 1))
row_its <- plot_grid(col_its_treatment, col_its_soil, col_its_incubation, 
                     ncol = 3, labels = "B.", label_size = 20, label_x =  -0.01, label_y = 1.02, hjust = -0.5, vjust = 2, rel_widths = c(1, 1, 1))
row_rps10 <- plot_grid(col_rps10_treatment, col_rps10_soil, col_rps10_incubation, 
                       ncol = 3, labels = "C.", label_size = 20, label_x =  -0.01, label_y = 1.02, hjust = -0.5, vjust = 2, rel_widths = c(1, 1, 1))


# Combine all elements with adjusted spacing
facetted_figure <- plot_grid(
  row_16S,
  row_its,
  row_rps10,
  nrow = 3,
  align = 'v',
  axis = 'l',
  rel_heights = c(1, 1, 1)  # Adjust relative heights
)

library(extrafont)
# Save the figure with increased dimensions
ggsave("Figures/facetted_figure_with_ordination_NMDS.svg", facetted_figure, width = 14, height = 14, units = "in")
ggsave("Figures/facetted_figure_with_ordination_NMDS.pdf", facetted_figure, width = 14, height = 14, units = "in")
```

### Software used
```{r software used}
sessioninfo::session_info()
```